---
title: "Hands-on Exercise 4: 2nd Order Spatial Point Patterns Analysis Methods"
date: "31 January 2023"
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

## Loading Packages

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

## Spatial Data Wrangling

### Importing the spatial data

```{r}
childcare_sf <- st_read("../../in-class_ex/in-class_ex04/data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "../../in-class_ex/in-class_ex04/data", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "../../in-class_ex/in-class_ex04/data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

DIY: Using the appropriate **sf** function you learned in Hands-on Exercise 2, retrieve the referencing system information of these geospatial data.

```{r}

```

DIY: Using the method you learned in Lesson 2, assign the correct crs to mpsz_sf and sg_sf simple feature data frames.

```{r}

```

DIY: If necessary, changing the referencing system to Singapore national projected coordinate system.

```{r}

```

### Mapping the geospatial data sets

DIY: Using the mapping methods you learned in Hands-on Exercise 3, prepare a map as shown below.

```{r}

```

```{r}
tmap_mode('view')
```

```{r}
tm_shape(childcare_sf)+
  tm_dots()
```

```{r}
tmap_mode('plot')
```

## Geospatial Data wrangling

### Converting sf data frames to sp's Spatial\* class

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

DIY: Using appropriate function, display the information of these three Spatial\* classes as shown below.

```{r}
childcare
```

```{r}
mpsz
```

```{r}
sg
```

```{r}

```

### Converting the Spatial\* class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r}
childcare_sp
```

```{r}
sg_sp
```

### Converting the generic sp format into spatstat's ppp format

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

### Handling duplicated points

```{r}
any(duplicated(childcare_ppp))
```

```{r}
multiplicity(childcare_ppp)
```

```{r}
sum(multiplicity(childcare_ppp) > 1)
```

```{r}
tmap_mode('view')
```

```{r}
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r}
tmap_mode('plot')
```

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

DIY: Using the method you learned in previous section, check if any dusplicated point in this geospatial data.

```{r}
any(duplicated(childcare_ppp_jit))
```

### Creating ***owin*** object

```{r}
sg_owin <- as(sg_sp, "owin")
```

```{r}
plot(sg_owin)
```

```{r}
summary(sg_owin)
```

### Combining point events object and owin object

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
```

```{r}
summary(childcareSG_ppp)
```

DIY: Using the method you learned in previous exercise, plot the newly derived childcareSG_ppp as shown below.

```{r}

```

#### Extracting study area

```{r}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```

Plotting target planning areas

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

#### Converting the spatial point data frame into generic sp format

```{r}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

#### Creating ***owin*** object

```{r}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

#### Combining childcare points and the study area

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

```{r}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

```{r}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

## Second-order Spatial Point Patterns Analysis

## Analysing Spatial Point Process Using G-Function
