---
title: "Take Home Exercise 1"
date: "9 February 2023"
date-modified: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

# 1.0 Overview

------------------------------------------------------------------------

## 1.1 Background

This analysis aims to apply appropriate spatial point patterns analysis methods to discover the geographical distribution of functional and non-function water points and their co-locations if any in Osun State, Nigeria.

## 1.2 Task

-   Exploratory Spatial Data Analysis (ESDA)

-   Second-order Spatial Point Pattern Analysis

-   Spatial Correlation Analysis

# 2.0 Setup

------------------------------------------------------------------------

## 2.1 Import Packages

-   sf - Used for handling geospatial data

-   tidyVerse - Used for data transformation and presentation

-   tmap, maptools, kableExtra - Used for visualizing dataframes and plots

-   spatstat - Used for point-pattern analysis

-   sfdep - Used for functions not in spdep

-   raster - Used to handle gridded spatial data

```{r}
pacman::p_load(sf, maptools, raster, spatstat, tmap, kableExtra, tidyverse, funModeling, sfdep)
```

# 3.0 Data Wrangling

------------------------------------------------------------------------

## 3.1 Datasets Used

```{r}
#| code-fold: true
# initialise a dataframe of our geospatial and aspatial dataset details
datasets <- data.frame(
  Type=c("Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         
         "Aspatial"),
  
  Name=c("geoBoundaries-NGA-ADM2",
         "geoBoundaries-NGA-ADM2",
         "geoBoundaries-NGA-ADM2",
         "geoBoundaries-NGA-ADM2",
         "geoBoundaries-NGA-ADM2",
         "geoBoundaries-NGA-ADM2",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         "nga_admbnda_adm2_osgof_20190417",
         
         "WPdx"),
  
  Format=c(".dbf", 
           ".geojson", 
           ".prj", 
           ".shp", 
           ".shx", 
           ".topojson",
           ".CPG",
           ".dbf",
           ".prj",
           ".sbn", 
           ".sbx", 
           ".shp", 
           ".shp", 
           ".shx", 
          
           ".csv"),
  
  Source=c("[geoBoundaries](https://www.geoboundaries.org/index.html#getdata)",
           "[geoBoundaries](https://www.geoboundaries.org/index.html#getdata)",
           "[geoBoundaries](https://www.geoboundaries.org/index.html#getdata)",
           "[geoBoundaries](https://www.geoboundaries.org/index.html#getdata)",
           "[geoBoundaries](https://www.geoboundaries.org/index.html#getdata)",
           "[geoBoundaries](https://www.geoboundaries.org/index.html#getdata)",
           
          "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           "[Humanitarian Data Exchange](https://data.humdata.org/dataset/cod-ab-nga)",
           
           "[ WPdx Global Data Repositories](https://www.waterpointdata.org/access-data/)")
  )

# with reference to this guide on kableExtra:
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# kable_material is the name of the kable theme
# 'hover' for to highlight row when hovering, 'scale_down' to adjust table to fit page width
library(knitr)
library(kableExtra)
kable(datasets, caption="Datasets Used") %>%
  kable_material("hover", latex_options="scale_down")

```

## 3.2 Geospatial Data

### 3.2.1 Load Data

Import Geospatial data and filter out Osun State. Transform WGS 48 coordinate system to projected coordinate system with Nigeria's ESPG - 26293

::: panel-tabset
#### Import

```{r}
osun <- st_read(dsn = "data/geospatial/",
               layer = "nga_admbnda_adm2_osgof_20190417") %>%
  filter(ADM1_EN == "Osun") %>%
  st_transform(crs = 26392)
```

#### Glimpse

```{r}
glimpse(osun)
```
:::

### 3.2.2 Data Preprocessing

#### 3.2.2.1 Exclude redundant fields

```{r}
osun <- osun %>%
  select(c(3:4, 8:9))
```

#### 3.2.2.2 Invalid Geometries

```{r}
length(which(st_is_valid(osun) == FALSE))
```

Everything is valid

#### 3.2.2.3 Checking for Duplicate Names

```{r}
osun$ADM2_EN[duplicated(osun$ADM2_EN)==TRUE]
```

No duplicate Local Government Areas (LGAs)

#### 3.2.2.4 Remove Missing Values

```{r}
osun[rowSums(is.na(osun))!=0,]
```

### 3.2.3 Initial Visualisation

```{r}
plot(st_geometry(osun), main="Osun State")
```

## 3.3 Aspatial Data

### 3.3.1 Load Data

Import Aspatial data and filter out Osun State.

::: panel-tabset
#### Import

```{r}
wp <- read_csv("data/aspatial/WPdx.csv") %>%
  filter(`#clean_country_name` == "Nigeria" & `#clean_adm1` == "Osun")
```

#### Glimpse

```{r}
glimpse(wp)
```
:::

### 3.3.2 Data Preprocessing

#### 3.3.2.1 Create sfc object column

Convert well known text (wkt) field into simple feature column (sfc) field

```{r}
wp$Geometry = st_as_sfc(wp$`New Georeferenced Column`)
wp$Geometry
```

#### 3.3.2.2 Create Simple Feature DataFrame

Store sfc as single feature (sf) object

```{r}
wp_sf <- st_sf(wp, crs=4326)
```

#### 3.3.2.3 Transform Projection System

```{r}
wp_sf <- wp_sf %>%
  st_transform(crs = 26392)
```

#### 3.3.2.4 Handle Missing Data

```{r}
freq(data = wp_sf,
     input = '#status_clean')
```

There are eight classes in the #status_clean fields.

To change NA with "unknown"

```{r}
wp_sf <- wp_sf %>% 
  rename(status_clean = '#status_clean') %>%
  select(status_clean) %>%
  mutate(status_clean = replace_na(
    status_clean, "unknown"))
```

### 3.3.3 Extract Water Points

Functional Points

```{r}
wp_f <- wp_sf %>%
  filter(status_clean %in%
           c("Functional",
             "Functional but needs repair",
             "Functional but not in use"))
```

Non-Functional Points

```{r}
wp_nf <- wp_sf %>%
  filter(status_clean %in%
           c("Non-Functional",
             "Non-Functional due to dry season",
             "Abandoned",
             "Abandoned/Decommissioned"))
```

Unknown Status Points

```{r}
wp_u <- wp_sf %>%
  filter(status_clean == "unknown")
```

::: panel-tabset
#### Functional Water Points

```{r}
freq(data = wp_f,
     input = 'status_clean')
```

#### Non-Functional Water Points

```{r}
freq(data = wp_nf,
     input = 'status_clean')
```

#### Unknown status Water Points

```{r}
freq(data = wp_u,
     input = 'status_clean')
```
:::

### 3.3.4 Point-in-Polygon Count

Extract number of total, functional, nonfunctional and unknown water points in each local government area (LGA)

```{r}
wp <- osun %>% 
  mutate(`total_wp` = lengths(st_intersects(osun, wp_sf))) %>%
  mutate(`wp_f` = lengths(st_intersects(osun, wp_f))) %>%
  mutate(`wp_nf` = lengths(st_intersects(osun, wp_nf))) %>%
  mutate(`wp_u` = lengths(st_intersects(osun, wp_u)))
```

To save as rds format

```{r}
#write_rds(wp, "Data/rds/wp.rds")
```

```{r}
ggplot(data = wp,
       aes(x = total_wp)) + 
  geom_histogram(bins=20,
                 color="black",
                 fill="light blue") +
  geom_vline(aes(xintercept=mean(
    total_wp, na.rm=T)),
             color="red", 
             linetype="dashed", 
             size=0.8) +
  ggtitle("Distribution of total water points by LGA") +
  xlab("No. of water points") +
  ylab("No. of LGAs") 
```

## 3.4 Combined Data Wrangling

### 3.4.1 Convert sf dataframes to sp Spatial\* class

```{r}
wp_f_spat = as_Spatial(wp_f)
wp_nf_spat = as_Spatial(wp_nf)
osun_spat = as_Spatial(osun)
```

Peek

::: panel-tabset
#### Functional Water Points

```{r}
wp_f_spat
```

#### Non-Functional Water Points

```{r}
wp_nf_spat
```

#### Osun State

```{r}
osun_spat
```
:::

### 3.4.2 Convert sp Spatial\* class to generic sp format

```{r}
wp_f_sp <- as(wp_f_spat, "SpatialPoints")
wp_nf_sp <- as(wp_nf_spat, "SpatialPoints")
osun_sp <-as(osun_spat, "SpatialPolygons")
```

Peek

::: panel-tabset
#### Functional Water Points

```{r}
wp_f_sp
```

#### Non-Functional Water Points

```{r}
wp_nf_sp
```

#### Osun State

```{r}
osun_sp
```
:::

### 3.4.3 Convert generic sp to spatstat ppp format

```{r}
# from sp object, convert into ppp format
wp_f_ppp <- as(wp_f_sp, "ppp")
wp_nf_ppp <- as(wp_nf_sp, "ppp")
```

View

```{r}
par(mfrow=c(1,2))
plot(wp_f_ppp, main="Functional Water Points")
plot(wp_nf_ppp, main="Non-functional Water Points")
```

#### 3.4.3.1 Check for Duplication

```{r}
any(duplicated(wp_f_ppp)); any(duplicated(wp_nf_ppp)) 
```

No duplicates ðŸ¤ no jittering

### 3.4.4 Create Owin object

```{r}
osun_owin <- as(osun_sp, "owin")
plot(osun_owin, main="Owin object of Osun State")
```

### 3.4.4 Combine point event object with Owin object

```{r}
wp_f_ppp = wp_f_ppp[osun_owin]
wp_nf_ppp = wp_nf_ppp[osun_owin]
```

View

```{r}
#| code-fold: true
par(mfrow=c(1,2))
plot(wp_f_ppp, main="Functional Water Points")
plot(wp_nf_ppp, main="Non-functional Water Points")
```

# 4.0 Exploratory Spatial Data Analysis (ESDA)

------------------------------------------------------------------------

## 4.1 Initial Visualization

::: panel-tabset
### Visual 1

```{r}
#| code-fold: true
wp_f_1 <- tm_shape(wp) +
  tm_fill("wp_f",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(title = "Functional Water Points by LGA",
            legend.outside = TRUE)

wp_f_2 <- tm_shape(wp) +
  tm_fill("wp_nf",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(title = "Non-functional Water Points by LGA",
            legend.outside = TRUE)


tmap_arrange(wp_f_1, wp_f_2, nrow = 1)
```

### Visual 2

```{r}
#| code-fold: true
tmap_mode("view") + 
tm_basemap("OpenStreetMap") +
tm_shape(wp_f) +
 tm_dots(col = "status_clean",
         pal = "blue",
         border.col = "black",
         title = "Functional") +
tm_shape(wp_nf) +
 tm_dots(col = "status_clean",
         pal = "purple",
         border.col = "black",
         title = "Non-Functional") +
  tm_view(set.bounds = c(4,7,5,8),
          set.zoom.limits = c(8,15)) 
```
:::

## 4.1 Kernel Density Estimation (KDE)

Sigma = `bw.ppl()` is used as there are multiple clusters and would be more prominent than `bw.diggle()`

::: panel-tabset
### Automatic Bandwidth

```{r}
kde_wp_f <- density(wp_f_ppp, 
                    sigma=bw.ppl, 
                    edge=TRUE,
                    kernel = "gaussian")

kde_wp_nf <- density(wp_nf_ppp,
                     sigma=bw.ppl,
                     edge=TRUE,
                     kernel = "gaussian")

par(mfrow=c(1,2))
plot(kde_wp_f, 
     main = "Functional Water Points (in m)",
     ribside=c("bottom"))
plot(kde_wp_nf,
     main = "Non-Functional Water Points (in m)",
     ribside=c("bottom"))
```

### Adaptive Bandwidth

```{r}
kde_wp_f_ad <- adaptive.density(wp_f_ppp,
                                  method = "kernel")

kde_wp_nf_ad <- adaptive.density(wp_nf_ppp,
                                     method = "kernel")

par(mfrow=c(1,2))
plot(kde_wp_f_ad,
     main = "Functional Water Points (in m)",
     ribside=c("bottom"))
plot(kde_wp_nf_ad,
     main = "Non-Functional Water Points (in m)",
     ribside=c("bottom"))
```
:::

### 4.2.1 Re-Scale KDE values

From m to km

```{r}
wp_f_ppp.km <- rescale(wp_f_ppp, 1000, "km")
wp_nf_ppp.km <- rescale(wp_nf_ppp, 1000, "km")
```

::: panel-tabset
#### Automatic Bandwidth

```{r}
kde_wp_f_scale <- density(wp_f_ppp.km,
                          sigma = bw.ppl,
                          edge = TRUE)

kde_wp_nf_scale <- density(wp_f_ppp.km,
                             sigma = bw.ppl,
                             edge = TRUE)

par(mfrow=c(1,2))
plot(kde_wp_f_scale,
     main = "Functional Water Points (in km)",
     ribside=c("bottom"))
plot(kde_wp_nf_scale,
     main = "Non-Functional Water Points (in km)",
     ribside=c("bottom"))
```

#### Adaptive Bandwidth

```{r}
kde_wp_f_scale_ad <- adaptive.density(wp_f_ppp.km,
                                  method = "kernel")

kde_wp_nf_scale_ad <- adaptive.density(wp_nf_ppp.km,
                                     method = "kernel")

par(mfrow=c(1,2))
plot(kde_wp_f_scale_ad,
     main = "Functional Water Points (in km)",
     ribside=c("bottom"))
plot(kde_wp_nf_scale_ad,
     main = "Non-Functional Water Points (in km)",
     ribside=c("bottom"))
```
:::

### 4.2.2 Convert KDE to Grid Object

::: panel-tabset
## Automatic Bandwidth

```{r}
grid_wp_f <- as.SpatialGridDataFrame.im(kde_wp_f_scale)
grid_wp_nf <- as.SpatialGridDataFrame.im(kde_wp_nf_scale)

spplot(grid_wp_f)
spplot(grid_wp_nf)
```

## Adaptive Bandwidth

```{r}
grid_wp_f_ad <- as.SpatialGridDataFrame.im(kde_wp_f_scale_ad)
grid_wp_nf_ad <- as.SpatialGridDataFrame.im(kde_wp_nf_scale_ad)

spplot(grid_wp_f_ad)
spplot(grid_wp_nf_ad)
```
:::

### 4.2.3 Convert grid object to Raster

::: panel-tabset
#### Functional Water Points

```{r}
kde_wp_f_raster <- raster(grid_wp_f)
kde_wp_f_raster
```

#### Non-Functional Water Points

```{r}
kde_wp_nf_raster <- raster(grid_wp_nf)
kde_wp_nf_raster
```
:::

### 4.2.4 Assign Projection System

::: panel-tabset
#### Functional Water Points

```{r}
projection(kde_wp_f_raster) <- CRS("+init=EPSG:26391 +datum:WGS84 +units=km")
kde_wp_f_raster
```

#### Non-Functional Water Points

```{r}
projection(kde_wp_nf_raster) <- CRS("+init=EPSG:26391 +datum:WGS84 +units=km")
kde_wp_nf_raster
```
:::

## 4.2 Visualize KDE

```{r}
#| code-fold: true
#tmap generation function

kde <- function(raster_obj, map_title) {
  tmap_mode("view")
  #tm_basemap("OpenStreetMap") +
tm_shape(raster_obj) +
  tm_raster("v", alpha=0.9, palette = "YlGnBu") +
  tm_view(set.bounds = c(4,7,5,8),
          set.zoom.limits = c(8, 13))
  } 
```

::: panel-tabset
### Functional Water Points

```{r}
kde(kde_wp_f_raster)
```

### Non-Functional Water Points

```{r}
kde(kde_wp_nf_raster)
```
:::

## 4.3 Nearest Neighbor Index

Clarke-Evans Test to check if distribution is random, clustered or dispersed with a 95% confidence interval

### 4.3.1 Test - Functional Water Points

The test hypotheses for Functional Water Point is :

**H~0~ :** The distribution is random

**H~1~ :** The distribution is not random

```{r}
clarkevans.test(wp_f_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("clustered"),
                nsim=99)
```

### 4.3.2 Test - Non-Functional Water Points

The test hypotheses for Non-Functional Water Point is :

**H~0~ :** The distribution is random

**H~1~ :** The distribution is not random (clustered)

```{r}
clarkevans.test(wp_nf_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("clustered"),
                nsim=99)
```

### 4.3.3 Conclusion

As the p-value is less than 0.05 is statistically significant and R\<1 for both functional and non-functional water points, we reject the null hypothesis that the distribution is not random and conclude that the **water points are likely clustered**.

# 5.0 2nd Order Spatial Point Pattern Analysis

Using G-Function for analysis with 95% confidence

------------------------------------------------------------------------

## 5.1 Test 1 - Functional Water Points

The test hypotheses for Functional Water Point is :

**H~0~ :** The distribution is random

**H~1~ :** The distribution is not random

```{r}
G_wp_f = Gest(wp_f_ppp, correction = "border")
plot(G_wp_f)
```

Complete Spatial Randomness Test

::: panel-tabset
#### Simulation

```{r}
G_wp_f.csr <- envelope(wp_f_ppp, Gest, nsim = 500)
```

#### Plot

```{r}
plot(G_wp_f.csr, main="G test - Functional")
```
:::

### 5.1.1 Conclusion

We reject the null hypothesis that functional water points are randomly distributed at 99% confidence interval. As G\^~obs~(r) rises quickly and is greater than G~theo~(r) and the envelope it means the functional water points are clustered.

## 5.2 Test 2 - Non-functional Water Points

The test hypotheses for Non-Functional Water Point is :

**H~0~ :** The distribution is random

**H~1~ :** The distribution is not random

```{r}
G_wp_nf = Gest(wp_nf_ppp, correction = "border")
plot(G_wp_nf)
```

Complete Spatial Randomness Test

::: panel-tabset
#### Simulation

```{r}
G_wp_nf.csr <- envelope(wp_nf_ppp, Gest, nsim = 500)
```

#### Plot

```{r}
plot(G_wp_nf.csr,  main = "G test - Non-functional")
```
:::

### 5.2.1 Conclusion

We reject the null hypothesis that non-functional water points are randomly distributed at 99% confidence interval. As G\^~obs~(r) rises quickly and is greater than G~theo~(r) and the envelope it means the functional water points are clustered.

## 5.3 Extra Tests

We see the same observation as mentioned above after zooming into a few neighborhoods such as Ife Central for functional water ponits and Ede North for non-functional water points.

### 5.3.1 Data Processing

```{r}
#Functional water points
IfeCentral <- osun[osun$ADM2_EN == "Ife Central",] %>%
    as('Spatial') %>%
    as('SpatialPolygons') %>%
    as('owin')


#non functional points
EdeNorth <- osun[osun$ADM2_EN == "Ede North",] %>%
    as('Spatial') %>%
    as('SpatialPolygons') %>%
    as('owin')


IfeCentral_ppp <- wp_f_ppp[IfeCentral]
EdeNorth_ppp <- wp_nf_ppp[EdeNorth]
```

View

```{r}
#| code-fold: true
par(mfrow=c(1,2))
plot(IfeCentral_ppp,
     main = "Ife Central - Functional Water Points",
     ribside=c("bottom"))
plot(EdeNorth_ppp,
     main = "Ede North - Non-Functional Water Points",
     ribside=c("bottom"))
```

### 5.3.2 Ife Central - Functional Water Points

```{r}
Ifecentral_G.csr = Gest(IfeCentral_ppp, correction = "border")
plot(Ifecentral_G.csr)
```

Complete Spatial Randomness Test

::: panel-tabset
#### Simulation

```{r}
Ifecentral_G.csr <- envelope(IfeCentral_ppp, Gest, nsim=100)
```

#### Plot

```{r}
plot(Ifecentral_G.csr, main="Ife Central G-Function Test")
```
:::

### 5.3.3 Ede North - Non-Functional Water Points

```{r}
EdeNorth_G.csr = Gest(EdeNorth_ppp, correction = "border")
plot(EdeNorth_G.csr)
```

Complete Spatial Randomness Test

::: panel-tabset
#### Simulation

```{r}
EdeNorth_G.csr <- envelope(EdeNorth_ppp, Gest, nsim=100)
```

#### Plot

```{r}
plot(EdeNorth_G.csr, main="Ede North G-Function Test")
```
:::

### 5.3.4 Conclusion 

We observe the same that G\^~obs~(r) rises quickly and is greater than G~theo~(r) and the envelope it means the functional and non-functional water points are clustered, thereby rejecting the null hypothesis that the water points are randomly distributed at 99% confidence interval.

# 6.0 Spatial Correlation Analysis

------------------------------------------------------------------------

## 6.1 Density Analysis

Segregate data for plotting. Then convert to spatstat ppp format

```{r}
wp_spat <- as_Spatial(wp_sf)
wp_spat@data$status_clean <-as.factor(wp_spat@data$status_clean)
wp_spat_ppp <- as(wp_spat, "ppp")
  
plot(wp_spat_ppp, which.marks = "status_clean")
```

## 6.3 Test Correlation

Use LCross function to test for spatial independence

The test hypotheses for the spatial distribution of functional and non-functional water point are:

**H~0~ :** The distribution is spatially independent

**H~1~ :** The distribution is not spatially independent

Simulation

```{r}
wp_spat_ppp_Lcross.csr <- envelope(wp_spat_ppp, 
                                 Lcross, 
                                 i="Functional", 
                                 j="Non-Functional", 
                                 correction="border", 
                                 nsim=100)
```

Plot

```{r}
plot(wp_spat_ppp_Lcross.csr, xlab="in km")
```

**Conclusion**

We reject the null hypothesis that water points are statistically independent at 99% confidence interval. As L\^~obs~(r) is greater than L~theo~(r) and the envelope it means the water points are statistically not independent.

## 6.4 Local Colocation Analysis

Remove status 'unknown' water points

```{r}
wp_sf_clean <- wp_sf %>%  filter(!status_clean=='unknown')
```

```{r}
#6 nearbest neighbors
nb = include_self(st_knn(st_geometry(wp_sf_clean), 6)) 

#weight matrix
wt = st_kernel_weights(nb, wp_sf_clean, "gaussian", adaptive = TRUE)


f = wp_sf_clean %>%
  filter(status_clean == "Functional")
A = f$status_clean

nf = wp_sf_clean %>%
  filter(status_clean == "Non-Functional")
B = nf$status_clean

LCLQ = local_colocation(A, B, nb, wt, 42)
LCLQ_wp = cbind(wp_sf_clean, LCLQ)
```

```{r}
LCLQ_wp <- LCLQ_wp %>% 
  mutate(
    `p_sim_Non.Functional` = replace(`p_sim_Non.Functional`, `p_sim_Non.Functional` > 0.05, NA),
    `Non.Functional` = ifelse(`p_sim_Non.Functional` > 0.05, NA, `Non.Functional`))

LCLQ_wp <- LCLQ_wp %>% mutate(`size` = ifelse(is.na(`Non.Functional`), 1, 5))
```

View

```{r}
#| code-fold: true
tmap_mode('view')
tm_view(set.bounds = c(4,7,5,8),
        set.zoom.limits=c(8, 13),
        bbox = st_bbox(filter(LCLQ_wp, !is.na(`Non.Functional`)))) +
  tm_shape(osun) +
  tm_borders() +
  tm_shape(LCLQ_wp) +
  tm_dots(col = c("Non.Functional"), 
          size = "size",
          scale=0.15,
          border.col = "black",
          border.lwd = 0.5,
          alpha=0.5,
          title="LCLQ"
          )

tmap_mode('plot')
```

LCLQ is just less than 1 which is much greater than the p-value of 0.05. Hence it means there are few Non-Functional Water Points in the same neighborhood as Functional Water Points.

# 7.0 Acknowledgements

------------------------------------------------------------------------

I'd like to thank Professor Kam for his insights and resource materials provided under IS415 Geospatial Analytics and Applications.
